---
title: "Hungary Chickenpox Case Analysis"
date: "5/3/2021"
output:
  pdf_document: default
  html_document: default
---


```{r}
library(MTS)
library(fpp)
library(TSA)
library(forecast)
library(Metrics)
library(fracdiff)
library(lubridate)
```


#load 52 weeks data
```{r}
dataPath<-"***"
```


```{r}
data<-read.csv(paste(dataPath,'hungary_chickenpox2.csv',sep = '/'), header=TRUE)
row.names(data)<-data$Date
Date<-data$Date
data<-data[-1]
```



```{r}
data$TOTAL<-rowSums(data)
```


```{r}
total<-ts(data$TOTAL,frequency=52,start=c(2005,1))
```


Plot total data
```{r}
plot(total)
```

KPSS test for total
```{r}
kpss.test(total)
```


Transformation
```{r}
#natural log
lambda<-0
trans.total<-BoxCox(total,lambda)
plot(trans.total)
```

KPSS test for trans total
```{r}
kpss.test(trans.total)
```



```{r}
diff<-diff(trans.total)
plot(diff,xlab="Years 2005-2014",main="Difference of Transformation Data", ylab="diff = 1")
```

```{r}
kpss.test(diff)
```


Split Data
```{r}
#Take train (2005-2013) 
train <- window(total,start=c(2005,1),end=c(2013,52))
train
```


```{r}
#test(2014)
test<-window(total,start=c(2014),end=c(2014,52))
test
```




```{r}
checkresiduals(train)
```

##1.Seasonal Arima model

```{r}
m1<-auto.arima(train,seasonal=TRUE,lambda="auto",trace=TRUE)
summary(m1)
```


```{r}
as.character(m1)
```


```{r}
tsdisplay(m1$residuals,main="Residuals from ARIMA(3,0,0)(1,1,0)[52] with drift")
```

```{r}
checkresiduals(m1,lag=56)
```

Plot forecast of SARIMA
```{r}
f1 <- forecast(m1, h = 52)$mean
plot(f1)
```

```{r}
autoplot(test,main="sARIMA Performance") +
  autolayer(f1, series="sARIMA") 
```


```{r}
mape1 <- mape(test, f1)
rmse1 <- rmse(test, f1)

cbind(mape1, rmse1)
```



```{r}
eacf(train)
```



```{r}
edaarima<-auto.arima(train,d=1,seasonal=TRUE,lambda=0,trace=TRUE)
```

```{r}
summary(edaarima)
checkresiduals(edaarima,lag=56)
```


##2.Seasonal naïve Model

```{r}

Model_SNaive <- snaive(train, 52)

```




```{r}
f3 <- forecast(Model_SNaive, h = 52)$mean

```



```{r}
autoplot(test) +
  autolayer(f3, series="Seasonal naïve")
  
```

```{r}
mape3 <- mape(test, f3)
rmse3 <- rmse(test, f3)

```


```{r}
tsdisplay(residuals(Model_SNaive),main ="Residuals from Seasonal naïve Model")
```

```{r}
checkresiduals(Model_SNaive,lag=56)
```
```{r}
summary(Model_SNaive)
```



##3.ARFIMA Model
```{r}
m2<-arfima(train,seasonal=TRUE)
summary(m2)
f4<-forecast(m2,h=52)$mean
```


```{r}
m2$hessian.dpq
```





```{r}
checkresiduals(m2,lag=56)
```

```{r}
Box.test(m2$residuals,type = c( "Ljung-Box"))
```



```{r}
rmse(train,m2$fitted)
mape(train,m2$fitted)
```


Plot forecast of ARFIMA
```{r}
autoplot(test,main="ARFIMA Performance") +
  autolayer(f4, series="ARFIMA")
```

```{r}

mape2 <- mape(test, f4)
rmse2 <- rmse(test, f4)

cbind(mape2, rmse2)
```


```{r}
tsdisplay(residuals(m2),main="ARFIMA")
```


##4.TBATS Model

```{r}
m3<-tbats(train,use.trend=TRUE)
summary(m3)
as.character(m3)
```


```{r}
rmse(train,m3$fitted.values)
mape(train,m3$fitted.values)
```


Plot forecast of TBATs

```{r}
foretbats<-forecast(m3,h=52)$mean

autoplot(test,main="TBATs Performance") +
  autolayer(foretbats, series="TBATs")



```



```{r}
tsdisplay(residuals(m3),main="Residuals from TBATS(0.005, {0,0}, 0.908, {<52,13>})")
```

```{r}
checkresiduals(m3,lag=56)
```


```{r}
mape4 <- mape(test, foretbats)
rmse4 <- rmse(test, foretbats)

cbind(mape4, rmse4)
```


##5.Fourier Model

```{r}
fitall<-list()
for (i in 1:25){
  fit <- auto.arima(train, xreg = fourier(train, K = i), seasonal = FALSE)
  fitall<-c(fitall,fit$aicc)
}


```


Potential k values
8 9 10 15 18 22

```{r}
fitall
```




```{r}
bestfit <- list(aicc=Inf)
for (i in 8:25){
  fit <- auto.arima(train, xreg = fourier(train, K = i), seasonal = FALSE)
  if (fit$aicc < bestfit$aicc)
    bestfit <- fit
  else break;
}
cat(i)

```

```{r}
summary(bestfit)
```


```{r}
checkresiduals(bestfit,lag=56)
```


Plot forecast of Fourier

```{r}
fc <- forecast(bestfit, xreg = fourier(train, K = 12, h = 52))

plot(fc)
```



```{r}
autoplot(test,main='Fourier Arima Performance') + autolayer(fc$mean,series="Fourier")
```

```{r}
mape5 <- mape(test, fc$mean)
rmse5 <- rmse(test, fc$mean)

cbind(mape5, rmse5)
```


##6.tslm model

```{r}
t1 <- tslm(train~trend + season)
summary(t1)
```


```{r}
rmse(train,t1$fitted.values)
mape(train,t1$fitted.values)
```


Plot forecast of tslm

```{r}
foretslm<-forecast(t1,h=52)$mean

autoplot(test) +
  autolayer(foretslm, series="tslm")



```


```{r}
tsdisplay(residuals(t1),main="Residuals from tslm")
```

```{r}
checkresiduals(t1,lag=56)
```


```{r}
mape7 <- mape(test, foretslm)
rmse7 <- rmse(test, foretslm)

cbind(mape7, rmse7)
```


Forecast of all models

```{r}
autoplot(test,main = "Model Performance") +
  autolayer(f1, series="sArima") +
  autolayer(foretbats, series="TBATs")+
  autolayer(f4,series="ARFIMA")+
  autolayer(f3, series="Seasonal naïve")+
  autolayer(fc$mean, series="Fourier")
  autolayer(foretslm,series="tslm")
```

Forecast of selected models

```{r}
autoplot(test,main="Selected Model Performance") +
  autolayer(f1, series="sArima") +
  autolayer(foretbats, series="TBATs")+
  autolayer(f4,series="ARFIMA")
```


Test rmse ans mape for all models

```{r}
data.frame('Metrics' = c('MAPE','RMSE'),
           'sArima' = c(mape1,rmse1),
           'ARFIMA'=c(mape2,rmse2),
           'Seasonal naïve' = c(mape3,rmse3),
           'TBATs' = c(mape4,rmse4),
           'Fourier'=c(mape5,rmse5),
           "tslm"=c(mape7,rmse7))
```




Load original data

```{r}
data2<-read.csv(paste(dataPath,'hungary_chickenpox.csv',sep = '/'), header=TRUE,stringsAsFactors = FALSE)
country<-rowSums(data2[,-1])
country_weekly<-ts(country,frequency = 365.25/7, start = decimal_date(dmy(data2[1,1])))
```

```{r}
country_weekly
```



```{r}
plot(country_weekly)
```


Make train from 2005-2012 and test from 2012-2013

```{r}
train2<-window(country_weekly,start=2005,end=2013)
test2<-window(country_weekly,start=2013,end=2014)
```


## TBATs Model no. 2

```{r}
m3tt<-tbats(train2)
summary(m3tt)
as.character(m3tt)
```


```{r}
foretbatstt<-forecast(m3tt,h=52)$mean

autoplot(test2,main="TBATs 13 Performance") +
  autolayer(foretbatstt, series="TBATs 13")
```

```{r}
tsdisplay(residuals(m3tt),main="Residuals from TBATS(0.01, {0,0}, 0.937, {<52.18,6>})")
```


```{r}
mape6 <- mape(test2, foretbatstt)
rmse6 <- rmse(test2, foretbatstt)

cbind(mape6, rmse6)
```

```{r}
checkresiduals(m3tt)
```


##Fourier Model no.2

```{r}
bestfit13 <- list(aicc=Inf)
for (i in 17:25){
  fit <- auto.arima(train2, xreg = fourier(train2, K = i), seasonal = FALSE, lambda = 0)
  if (fit$aicc < bestfit13$aicc)
    bestfit13 <- fit
  else break;
}
# I = 12 or 15 or 18
```


```{r}
#summary(bestfit)
fc13 <- forecast(bestfit13, xreg = fourier(train2, K = 18, h = 52))
# fc2 <- forecast(bestfit, xreg = fourier(train_org, K = 2, h = 52), lambda = 0)
plot(fc13)
```

```{r}
autoplot(test2,main='Fourier Arima') + autolayer(fc13$mean)
```

```{r}
rmse(fc13$mean, test2)
mape(fc13$mean, test2)
```


##sArima model no.2

```{r}
m113<-auto.arima(train2,seasonal=TRUE,lambda="auto",trace=TRUE)
summary(m113)
as.character(m113)
```

```{r}
checkresiduals(m113)
```

```{r}
f113 <- forecast(m113, h = 52)$mean
plot(f113)
```


```{r}
autoplot(test2,main="sARIMA 13 Performance") +
  autolayer(f113, series="sARIMA 13") 
```

```{r}
rmse(test2,f113)
mape(test2,f113)
```

